### 2014-12-17 revision/Paper_v3.tex
final revisions

\subsection*{Estimating parameters and effective allosteric regulatory structures.}
A critical challenge for any dynamic model is the estimation of kinetic parameters. 
For metabolic processes, there is also the added challenge of identifying the regulation and control structures that manage metabolism. 
Of course, these issues are not independent; any description of enzyme activity regulation will be a function of system state, which in turn depends upon the kinetic parameters. 
For cell free systems, regulated gene expression has been removed, however, enzyme activity regulation is still operational. 
We explored this linkage by estimating model parameters from synthetic data using both network structures. 
We generated synthetic measurements of the substrate S, intermediate M$_{5}$ and end product P$_1$ approximately every 20 min using
network A. This data set is similar to published cell free studies both in terms of network coverage, and sampling frequency \citep{Jewett:2008aa}. 
We then generated an ensemble of model parameter estimates by minimizing the difference between model simulations and the synthetic data using particle swarm optimization (PSO), starting from random initial parameter guesses. The estimation of kinetic parameters was sensitive to the choice of regulatory structure (Fig. \ref{fig-parameter-fit}). 
PSO identified an ensemble of parameters that bracketed the mean of the synthetic measurements in less than 1000 iterations when the control structure was correct (Fig. \ref{fig-parameter-fit}A and B). 
However, with control mismatch (network B simulated with network A parameters), model simulations were not consistent with the synthetic data  (Fig. \ref{fig-parameter-fit}C and D). 
Taken together, these results suggested that we could perhaps simultaneously estimate both parameters and network control architectures, as incorrect control structures would be manifest as poor model fits.

@@

A critical challenge will be the estimation of allosteric connectivity at a genome scale. 
The regulation of glycolytic enzymes, such as phosphofructokinase I, has been studied for many years \citep{BERG_BIOCHEMISTRY,Peskov:2008aa}. 
The allosteric regulation of metabolic enzymes can also be established from organism specific databases, such as EcoCyc \citep{Keseler:2013aa}, 
or more general allosteric databases, such as the AlloSteric Database \citep{Huang:2014aa}. 
However, for those enzymes that have not been well studied, we will need to infer allosteric interactions from experimental data.
In general, the reverse engineering of regulatory network structure from data is a very difficult problem. 
Recently, Sauer and colleagues have developed a systematic, model-based approach for the identification of allosteric regulation \textit{in vivo} \cite{Link:2013aa}. 
They tested the effects of many putative allosteric protein-metabolite interactions on the performance of a kinetic model of glycolysis against dynamic metabolomic and fluxomic measurements. 
A method similar to this may be easily applied to cell-free systems in order to identify relevant \textit{in vitro} allosteric interactions. 
Because omics measurements of cell-free environments are easy to obtain, identification of large-scale allosteric control structures may be possible. 
Also, there are many different approaches from the reverse engineering of gene regulatory networks that perhaps could be adopted to this problem, however this remains an open question.
For example, one could imagine designing pulse chase experiments which maximally distinguish between competing allosteric models, similar to the earlier work of Kremling et al \citep{Kremling:2004aa}, or
iteratively estimate model structures similar to Doyle and coworkers \citep{Gadkar:2005aa}.
Lastly, the choice of max/min integration rules or the particular form of the transfer functions could be generalized to include other rule types and functions. 
Theoretically, an integration rule is a function whose domain is a set of transfer function inputs, and whose range is $v\in[0,1]$.
Thus, integration rules other than max/min could be used, such as the mean or the product, assuming the range of the transfer functions is always $f\in[0,1]$.
Alternative integration rules such as the mean might have different properties which could influence model identification or performance. 
For example, a mean integration rule would be differentiable, which allows derivative-based optimization approaches to be used. 
The particular form of the transfer function could also be explored. We choose a Hill-like function because of its 
prominence in the systems and synthetic biology community. However, the only mathematical requirement for a transfer function is that it map a non-negative continuous or categorical variable into the 
range $f\in[0,1]$. Thus, many types of transfer functions are possible.

@@

We minimized the model residual using particle swarm optimization (PSO) \citep{PSO}.
PSO uses a \textit{swarming} metaheuristic to explore parameter spaces. 
A strength of PSO is its ability to find the global minimum, even in the presence of potentially many local minima, by communicating the local
error landscape experienced by each particle collectively to the swarm. Thus, PSO acts both as a local and a global search algorithm. 
For each iteration, particles in the swarm compute their local error by evaluating the model equations using their specific parameter vector realization.
From each of these local points, a globally best error is identified. Both the local and global error 
are then used to update the parameter estimates of each particle using the rules:
\begin{eqnarray}
	\mathbf{\Delta}_{i} &=&\theta_{1}\mathbf{\Delta}_{i} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{k}_{i}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G} - \mathbf{k}_{i}\right) \\
	\mathbf{k}_{i} &=& \mathbf{k}_{i} + \mathbf{\Delta}_{i}
\end{eqnarray}
where $\mathbf{\Delta_i}$ denotes the perturbation to the vector of parameters $\mathbf{k_i}$ for particle $i$. 
$\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best local solution found by particle $i$, and
$\mathcal{G}$ denotes the best solution found over the entire population of particles. The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
parameters ($\mathcal{K}\times{1}$). In this study, we used $\left(\theta_{1},\theta_{2},\theta_{3}\right) = \left(1.0, 0.05564, 0.02886\right)$. The quality of parameter
estimates was measured using two criteria, goodness of fit (model residual) and angle between the estimated parameter vector $\mathbf{k}_{j}$ and the true parameter set $\mathbf{k}^{*}$:
\begin{equation}
	\alpha_{j} = cos^{-1}\left(\frac{\mathbf{k}_{j}\cdot{\mathbf{k^{*}}}}{\norm{\mathbf{k}_{j}}\norm{\mathbf{k^{*}}}}\right)
\end{equation}If the candidate parameter set $\mathbf{k}_{j}$ were perfect, the residual between the model and synthetic data 
and the angle between $\mathbf{k}_{j}$ and the true parameter set $\mathbf{k}^{*}$ would be equal to zero.

>>>>

\subsection*{Estimating parameters and effective allosteric regulatory structures.}
A critical challenge for any dynamic model is the estimation of kinetic parameters. 
For metabolic processes, there is also the added challenge of identifying the regulation and control structures that manage metabolism. 
Of course, these issues are not independent; any description of enzyme activity regulation will be a function of system state, which in turn depends upon the kinetic parameters. 
For cell-free systems, regulated gene expression has been removed, however, enzyme activity regulation is still operational. 
We explored this linkage by estimating model parameters from synthetic data using both network structures. 
We generated synthetic measurements of the substrate S, intermediate M$_{5}$ and end product P$_1$ approximately every 20 min using
network A. This data set is similar to published cell-free studies, both in terms of network coverage and sampling frequency \citep{Jewett:2008aa}. 
We then generated an ensemble of model parameter estimates by minimizing the difference between model simulations and the synthetic data using particle swarm optimization (PSO), starting from random initial parameter guesses. The estimation of kinetic parameters was sensitive to the choice of regulatory structure (Fig. \ref{fig-parameter-fit}). 
PSO identified an ensemble of parameters that bracketed the mean of the synthetic measurements in less than 1000 iterations when the control structure was correct (Fig. \ref{fig-parameter-fit}A and B). 
However, with control mismatch (network B simulated with network A parameters), model simulations were not consistent with the synthetic data  (Fig. \ref{fig-parameter-fit}C and D). 
Taken together, these results suggested that we could perhaps simultaneously estimate both parameters and network control architectures, as incorrect control structures would be manifest as poor model fits.

@@

A critical challenge will be the estimation of allosteric connectivity at a genome scale. 
The regulation of glycolytic enzymes, such as phosphofructokinase I, has been studied for many years \citep{BERG_BIOCHEMISTRY,Peskov:2008aa}. 
The allosteric regulation of metabolic enzymes can also be established from organism specific databases, such as EcoCyc \citep{Keseler:2013aa}, 
or more general allosteric databases, such as the AlloSteric Database \citep{Huang:2014aa}. 
However, for those enzymes that have not been well studied, we will need to infer allosteric interactions from experimental data.
In general, the reverse engineering of regulatory network structure from data is a very difficult problem. 
Recently, Sauer and colleagues have developed a systematic, model-based approach for the identification of allosteric regulation \textit{in vivo} \cite{Link:2013aa}. 
They tested the effects of many putative allosteric protein-metabolite interactions on the performance of a kinetic model of glycolysis against dynamic metabolomic and fluxomic measurements. 
A method similar to this may be easily applied to cell-free systems in order to identify relevant \textit{in vitro} allosteric interactions. 
Because omics measurements of cell-free environments are easy to obtain, identification of large-scale allosteric control structures may be possible. 
Also, there are many different approaches from the reverse engineering of gene regulatory networks that perhaps could be adopted to this problem, however this remains an open question.
For example, one could imagine designing pulse chase experiments which maximally distinguish between competing allosteric models, similar to the earlier work of Kremling et al \citep{Kremling:2004aa}, or
iteratively estimating model structures similar to Doyle and coworkers \citep{Gadkar:2005aa}.
Lastly, the choice of max/min integration rules or the particular form of the transfer functions could be generalized to include other rule types and functions. 
Theoretically, an integration rule is a function whose domain is a set of transfer function inputs, and whose range is $v\in[0,1]$.
Thus, integration rules other than max/min could be used, such as the mean or the product, assuming the range of the transfer functions is always $f\in[0,1]$.
Alternative integration rules such as the mean might have different properties which could influence model identification or performance. 
For example, a mean integration rule would be differentiable, which allows derivative-based optimization approaches to be used. 
The particular form of the transfer function could also be explored. We choose a Hill-like function because of its 
prominence in the systems and synthetic biology community. However, the only mathematical requirement for a transfer function is that it map a non-negative continuous or categorical variable into the 
range $f\in[0,1]$. Thus, many types of transfer functions are possible.

@@

We minimized the model residual using particle swarm optimization (PSO) \citep{PSO}.
PSO uses a \textit{swarming} metaheuristic to explore parameter spaces. 
A strength of PSO is its ability to find the global minimum, even in the presence of potentially many local minima, by communicating the local
error landscape experienced by each particle collectively to the swarm. Thus, PSO acts both as a local and a global search algorithm. 
For each iteration, particles in the swarm compute their local error by evaluating the model equations using their specific parameter vector realization.
From each of these local points, a globally best error is identified. Both the local and global error 
are then used to update the parameter estimates of each particle using the rules:
\begin{eqnarray}
	\mathbf{\Delta}_{i} &=&\theta_{1}\mathbf{\Delta}_{i} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{k}_{i}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G} - \mathbf{k}_{i}\right) \\
	\mathbf{k}_{i} &=& \mathbf{k}_{i} + \mathbf{\Delta}_{i}
\end{eqnarray}
where $\mathbf{\Delta}_i$ denotes the perturbation to the vector of parameters $\mathbf{k}_i$ for particle $i$. 
$\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best local solution found by particle $i$, and
$\mathcal{G}$ denotes the best solution found over the entire population of particles. The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
parameters ($\mathcal{K}\times{1}$). In this study, we used $\left(\theta_{1},\theta_{2},\theta_{3}\right) = \left(1.0, 0.05564, 0.02886\right)$. The quality of parameter
estimates was measured using two criteria, goodness of fit (model residual) and angle between the estimated parameter vector $\mathbf{k}_{j}$ and the true parameter set $\mathbf{k}^{*}$:
\begin{equation}
	\alpha_{j} = cos^{-1}\left(\frac{\mathbf{k}_{j}\cdot{\mathbf{k^{*}}}}{\norm{\mathbf{k}_{j}}\norm{\mathbf{k^{*}}}}\right)
\end{equation}If the candidate parameter set $\mathbf{k}_{j}$ were perfect, the residual between the model and synthetic data 
and the angle between $\mathbf{k}_{j}$ and the true parameter set $\mathbf{k}^{*}$ would be equal to zero.
